---
title: "Basic Plotting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Plotting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

First load libraries.

```{r setup}
library(dataviz)
library(ggplot2)
```

Examine data

```{r}
data_ala
```

* Use the column names to specify what goes on the `y` axis and what goes on the `x` axis.
* Use `stat_summary` to say that we don't want to plot the raw data, we want a summary statistic instead.
* Use `geom` to say that we want a bar chart, and `fun` to say that we want the mean (as opposed to the median, say). 

```{r}
ggplot(data_ala, aes(y=RT, x=Valence))+
  stat_summary(geom="bar", fun="mean")
```

We can add error bars

```{r}
ggplot(data_ala, aes(y=RT, x=Valence))+
  stat_summary(geom="bar", fun="mean")+
  stat_summary(geom='errorbar', width=0.3, fun.data = mean_cl_normal, fun.args=list(conf.int=.95))
```

We can use colour to distinguish the levels of Valence.

```{r}
ggplot(data_ala, aes(y=RT, x=Valence, colour=Valence, fill=Valence))+
  stat_summary(geom="bar", fun="mean", alpha=0.5)+
  stat_summary(geom='errorbar', width=0.3, fun.data = mean_cl_normal, 
               fun.args=list(conf.int=.95), show.legend = F)
```

